### 2020-05-27 writeup/proof_main.tex
Update on Overleaf.

We now state the formal version of Theorem \ref{thm_main_informal}. First, we need to introduce several quantities that will be used in our statement.

@@

\begin{proof}[Proof of Theorem \ref{thm_model_shift}]
\noindent\todo{A proof outline; including the following key lemma.}
To prove Theorem \ref{thm_cov_shift}, we study the spectrum of the random matrix model:
$$Q= \Sigma_1^{1/2}  Z_1^{\top} Z_1 \Sigma_1^{1/2}  + \Sigma_2^{1/2}  Z_2^{\top} Z_2 \Sigma_2^{1/2} ,$$
where $\Sigma_{1,2}$ are $p\times p$ deterministic covariance matrices, and $X_1=(x_{ij})_{1\le i \le n_1, 1\le j \le p}$ and $X_2=(x_{ij})_{n_1+1\le i \le n_1+n_2, 1\le j \le p}$ are $n_1\times p$ and $n_2 \times p$ random matrices, respectively, where the entries $x_{ij}$, $1 \leq i \leq n_1+n_2\equiv n$, $1 \leq j \leq p$, are real independent random variables satisfying
\begin{equation}\label{eq_12moment} %\label{assm1}
\mathbb{E} z_{ij} =0, \ \quad \ \mathbb{E} \vert z_{ij} \vert^2  = 1.
\end{equation}

The proof of Theorem \ref{thm_model_shift} involves two parts.

@@

\begin{lemma}\label{lem_cov_derivative}
	In the setting of Theorem \ref{thm_model_shift}, we have with high probability $1-{\rm o}(1)$,
\begin{equation}\label{lem_cov_derv_eq}
\begin{split}
&\wh w^{2}(n_1+n_2)^2\bignorm{\Sigma_2^{1/2} (\hat{w}X_1^{\top}X_1 + X_2^{\top}X_2)^{-1}\Sigma_1 (\beta_s - {w}\beta_t)}^2 \\
&= (\beta_s - {w}\beta_t)^{\top} \Sigma_1^{1/2} {M} \frac{(1 + a_3)\id + a_4 {M}^{\top}{M}}{(a_2 + a_1 {M}^{\top}{M})^2} {M}^{\top} \Sigma_1^{1/2} (\beta_s - \hat{w}\beta_t) +\OO(n^{-1/2+\e}),
\end{split}
\end{equation}
	for any constant $\epsilon>0$. %, where $a_{3,4}$ are found using equations in  \eqref{m35reduced}
\end{lemma}
We will give the proof of this lemma in Section \ref{sec_maintools}.
\end{proof}

@@

{\cor add some arguments with $\e$-net.}

>>>>

In this section, we state and prove the formal version of Theorem \ref{thm_main_informal}. First, we need to introduce several quantities that will be used in our statement.

@@

Now we first give a proof of Theorem \ref{thm_model_shift} based on Lemma \ref{lem_cov_shift}.

\begin{proof}[Proof of Theorem \ref{thm_model_shift}]
%\noindent
%To prove Theorem \ref{thm_cov_shift}, we study the spectrum of the random matrix model:
%$$Q= \Sigma_1^{1/2}  Z_1^{\top} Z_1 \Sigma_1^{1/2}  + \Sigma_2^{1/2}  Z_2^{\top} Z_2 \Sigma_2^{1/2} ,$$
%where $\Sigma_{1,2}$ are $p\times p$ deterministic covariance matrices, and $X_1=(x_{ij})_{1\le i \le n_1, 1\le j \le p}$ and $X_2=(x_{ij})_{n_1+1\le i \le n_1+n_2, 1\le j \le p}$ are $n_1\times p$ and $n_2 \times p$ random matrices, respectively, where the entries $x_{ij}$, $1 \leq i \leq n_1+n_2\equiv n$, $1 \leq j \leq p$, are real independent random variables satisfying
%\begin{equation}\label{eq_12moment} %\label{assm1}
%\mathbb{E} z_{ij} =0, \ \quad \ \mathbb{E} \vert z_{ij} \vert^2  = 1.
%\end{equation}
\todo{A proof outline; including the following key lemma.}

The proof is divided into four parts.

@@

\begin{lemma}\label{lem_cov_derivative}
	In the setting of Theorem \ref{thm_model_shift}, we have with high probability $1-{\rm o}(1)$,
\begin{equation}\label{lem_cov_derv_eq}
\begin{split}
&\wh w^{2}(n_1+n_2)^2\bignorm{\Sigma_2^{1/2} (\hat{w}X_1^{\top}X_1 + X_2^{\top}X_2)^{-1}\Sigma_1 (\beta_s - {w}\beta_t)}^2 \\
&= (\beta_s - {w}\beta_t)^{\top} \Sigma_1^{1/2} {M} \frac{(1 + a_3)\id + a_4 {M}^{\top}{M}}{(a_2 + a_1 {M}^{\top}{M})^2} {M}^{\top} \Sigma_1^{1/2} (\beta_s - \hat{w}\beta_t) +\OO(n^{-1/2+\e}),
\end{split}
\end{equation}
	for any constant $\epsilon>0$. %, where $a_{3,4}$ are found using equations in  \eqref{m35reduced}
\end{lemma}
We will give the proof of this lemma in Section \ref{sec_maintools}.

\paragraph{Part III: An $\e$-net argument.}

{\cor add some arguments with $\e$-net.}
\end{proof}

{\cor add some remarks about the difficulty in random matrix theory}

By applying a tighter bound to Part I of the above proof, we can conclude the proof Lemma \ref{prop_model_shift_tight}.

@@

### 2020-05-27 writeup/transfer.tex
merge?
Update on Overleaf.

\section{Revisiting the Bias-Variance Trade-off for Multi-Task learning}
\label{sec_insight}

We revisit the classical bias-variance trade-off in the context of multi-task learning for linear regressions, in particular the setup of equation \eqref{eq_mtl}.
For the high-dimensional setting where the data size of the target task is a small constant times dimension $p$, we quantify the trade-off as a function of the task models $\set{\beta_i}_{i=1}^t$, data size $\set{n_i}_{i=1}^t$, and covariance matrices $\set{\Sigma_i}_{i=1}^t$.
After describing this technical result, we provide theoretical insights over when multi-task learning performs better than single-task learning.
Our analysis reveals three contributing causes of negative transfer, including task model \textit{dissimilarity}, \textit{unbalanced} data size, and \textit{covariate shift}, which we describe in detail for the rest of the section.

\subsection{Determining Positive versus Negative Transfer}\label{sec_main}

In order to study the trade-off between model-shift bias and variance reduction, we need tight concentration bounds to quantify both effects.
For this purpose, we consider the high-dimensional setting where $p$ goes to infinity.
A well-known result for this setting states that the asymptotic limit of $\te(\hat{\beta}_t^{\STL})$ is $\frac {\sigma^2} {\rho_t - 1}$ (e.g. Chapter 6 of \cite{S07}), which scales with the data size and noise level of the target task.
However, this result only applies to the single-task setting.
Therefore, we focus on deriving the asymptotic limit of $\te(\hat{\beta}_t^{\MTL})$ when $p$ goes to infinity.

Our first result applies to the setting of two tasks where their covariance matrices may be arbitrarily different.
%The technical crux of our approach is to derive the asymptotic limit of $\te(\hat{\beta}_t^{\MTL})$ in the high-dimensional setting, when $p$ approaches infinity.
We derive a precise limit of $\bigtr{(X_1^{\top}X_1 + X_2^{\top}X_2)^{-1}\Sigma_2}$, which is a deterministic function that only depends on $\Sigma_1, \Sigma_2$ and $n_1/p, n_2/p$ (see Lemma \ref{lem_cov_shift} in Appendix \ref{app_proof_main} for the result).
Based on the result, we show how to determine positive versus negative transfer as follows.

@@

@@

\subsection{Task Model Similarity}\label{sec_similarity}

@@

\begin{example}\label{ex_basic}
	Consider a setting where $\Sigma_1 = \Sigma_2 = \id$, in other words there is no covariate shift between the two tasks. For the noises, we assume that $\e_i$ has i.i.d. entries with mean zero and variances $\sigma_i^2$, $i=1,2$. 
	For the task models, suppose that $\beta_2$ has i.i.d. entries with mean zero and variance $\kappa^2$ and $\beta_1 - \beta_2$ also has i.i.d. entries with mean $0$ and variance $d^2$.
	%The noise level for both tasks is equal to $\sigma > 0$.
	We have $n_i = c_i \cdot p$ data points from each task, for $i = 1, 2$.
	For simplicity, we assume that all the random variables have subexponential decay, while keeping in mind that our results can be applied under weaker moments assumptions as shown in the supplementary material.
\end{example}

@@

\subsection{Data Size}\label{sec_data_size}

@@

\subsection{Covariate Shift}\label{sec_covariate}

@@

\section{Extending the Bias-Variance Trade-off to Transfer Learning}\label{sec_taskonomy}

While our focus is on multi-task learning, we show that our results extend to transfer learning settings naturally.
We provide an analysis of the transfer function of Taskonomy \cite{ZSSGM18} using our setup.
Specifically, the source task encoder consists of the representations learnt from one or more source tasks.
The transfer function then tries to fit the target task data to the source task encoder.
For more details, we refer the reader to Figure 4 in Taskonomy.

>>>>

\section{A Technical Tool to Quantify the Trade-off}
\label{sec_main}

We revisit the classical bias-variance trade-off in the context of multi-task learning for linear regressions, in particular the setup of equation \eqref{eq_mtl}.
For the high-dimensional setting where the data size of the target task is a small constant times dimension $p$, we quantify the trade-off as a function of the task models $\set{\beta_i}_{i=1}^t$, data size $\set{n_i}_{i=1}^t$, and covariance matrices $\set{\Sigma_i}_{i=1}^t$.

\textbf{Multi-task learning.} Our first result applies to the setting of two tasks where their covariance matrices may be arbitrarily different.
%The technical crux of our approach is to derive the asymptotic limit of $\te(\hat{\beta}_t^{\MTL})$ in the high-dimensional setting, when $p$ approaches infinity.
We derive a precise limit of $\bigtr{(X_1^{\top}X_1 + X_2^{\top}X_2)^{-1}\Sigma_2}$, which is a deterministic function that only depends on $\Sigma_1, \Sigma_2$ and $n_1/p, n_2/p$ (see Lemma \ref{lem_cov_shift} in Appendix \ref{app_proof_main} for the result).
Based on the result, we show how to determine positive versus negative transfer as follows.

@@

\textit{Proof overview.} By comparing the bias and variance of $\te_t(\hat{\beta}_t^{\MTL})$ to $\te_t(\hat{\beta}_t^{\STL})$, we observe the following two effects, respectively.
%We begin by observing that the test error of $\hat{\beta}_t^{\MTL}$ consists of two parts.
We term equation \eqref{eq_te_model_shift} as \textit{model-shift bias}, which captures how similar $\beta_1$ and $\beta_2$ are.
This part introduces a negative effect to multi-task learning.
The scaling term $\hat{v}$ corresponds to the intuition that the shared module $B$ learns a subspace for both tasks.
The output layer of each task scales the direction of $B$ suitably to fit the task.

Next, it is not hard to verify that equation \eqref{eq_te_var}, the variance of $\hat{\beta}_t^{\MTL}$, is always smaller than the variance of $\hat{\beta}_t^{\STL}$.
This part introduces a positive variance reduction effect to performing multi-task learning.
Hence, whether $\te(\hat{\beta}_t^{\MTL})$ is lower than $\te(\hat{\beta}_t^{\STL})$ is determined by the trade-off between two effects:
(i) the positive effect from variance reduction;
(ii) the negative effect from model shift bias.

@@

\textbf{Extensions to transfer learning.}
While our focus is on multi-task learning, we show that our results extend to transfer learning settings naturally.
We provide an analysis of the transfer function of Taskonomy \cite{ZSSGM18} using our setup.
Specifically, the source task encoder consists of the representations learnt from one or more source tasks.
The transfer function then tries to fit the target task data to the source task encoder.
For more details, we refer the reader to Figure 4 in Taskonomy.

\section{Theoretical Implications}
\label{sec_insight}

After describing this technical result, we provide theoretical insights over when multi-task learning performs better than single-task learning.
Our analysis reveals three contributing causes of negative transfer, including task model \textit{dissimilarity}, \textit{unbalanced} data size, and \textit{covariate shift}, which we describe in detail for the rest of the section.

\subsection{De-noising by Transferring from a Source Task}\label{sec_similarity}

@@

\textit{A simplifying setting.}\label{ex_basic}
	Consider a setting where $\Sigma_1 = \Sigma_2 = \id$, in other words there is no covariate shift between the two tasks. For the noises, we assume that $\e_i$ has i.i.d. entries with mean zero and variances $\sigma_i^2$, $i=1,2$. 
	For the task models, suppose that $\beta_2$ has i.i.d. entries with mean zero and variance $\kappa^2$ and $\beta_1 - \beta_2$ also has i.i.d. entries with mean $0$ and variance $d^2$.
	%The noise level for both tasks is equal to $\sigma > 0$.
	We have $n_i = c_i \cdot p$ data points from each task, for $i = 1, 2$.
	For simplicity, we assume that all the random variables have subexponential decay, while keeping in mind that our results can be applied under weaker moments assumptions as shown in the supplementary material.

@@

\subsection{Improving Labeled Data Efficiency}\label{sec_data_size}

@@

\subsection{When to Align Task Covariances under Covariate Shift}\label{sec_covariate}

@@

%\subsection{Extension to Transfer Learning}\label{sec_taskonomy}
